---
title: "Integrated analysis and modelling of contaminant mixtures and transcriptomic responses in Tree Swallow (Tachycineta bicolor) nestlings in the Great Lakes"
output:
  html_document:
    df_print: paged
    theme: cerulean
    code_download: true
---

Chi Yen Tseng^1^, Christine M. Custer^2^, Thomas W. Custer^2^, Paul M. Dummer^2^, Natalie Karouna‐Renier^3^ and Cole W. Matson^1^

1. Department of Environmental Science, The Institute of Ecological, Earth, and Environmental Sciences (TIE3S), and the Center for Reservoir and Aquatic Systems Research (CRASR), Baylor University, Waco, Texas 76798, United States 
2. Upper Midwest Environmental Sciences Center, U.S. Geological Survey, La Crosse, Wisconsin 54603, United States 
3. U.S. Geological Survey, Eastern Ecological Science Center (EESC) at Patuxent, Beltsville, Maryland 20705, United States


### Load packages 
```{r Load packages, eval=FALSE}
library(glmnet)
library(edgeR)
library(readr)
library(tidyverse)
library(foreach)
library(doParallel)
library(DESeq2)
```


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 150px;
}
```

## 1. For selecting top predictor genes to predcit PCBs (by individual nestlings)  

### Load all data 

`GLRI_coldata`  
: chemistry data ("Dioxin" "MultiRedsiduePest" "PAHs" "PBDEs" "LRPCBs" "HRPCBs" "Pesticides" "PFCs" "PPCPs")  

`data.contaminants.majorSubset.geo.bysite`
: geometric mean of each contaminant key by site  

`contaminants.coldata.subset`
: only includes the nestlings with genomic information and removes PAHs data  

`ID.convert`
: gene ID convertion table between tree swallow gene pseudo name and chicken gene name  

`siteMAP.all`
: site ID and propersite name convertion Table  

`dds.bothSex.adjusted or counts.tswallow`
: normalized count matrix for all nestlings, adjusted for batch difference  

`Cont.major.list`
: major contaminants ("Nonachlor, cis-_C", "Heptachlor Epoxide_C",   "Nonachlor, trans-_C",  "PFDA_C", "Chlordane, oxy-_C", "Total Parent PAHs", "Total PBDE", "Total PAHs","Total PCBs","Total_PFCs")

`counts.tswallow.norm`
: vst transformed normalized counts   

```{r Load data, eval=FALSE}
setwd("/home/chiyen/Documents/work/Tswallow_chem_GLRI_update/GLRI_MS2_all")
# load chemistry data and show class
GLRI_coldata <- readRDS("GLRI_coldata_to2020.rds")
print(c("chemistry data class:", unique(GLRI_coldata$class)))
# Determine min contamin value for each Key
GLRI_coldata <- GLRI_coldata$contaminants %>% filter(!is.na(Value))
# Load chemistry GeoMean by site 
data.contaminants.majorSubset.geo.bysite <- readRDS("data_contaminants_majorSubset_geo_bysite.rds")
# Determine minimum conc. for each chemical 
contaminants.min.adjusted <- GLRI_coldata %>% group_by(Key,class) %>% summarise(minValue = min(as.numeric(value.adjust),na.rm = TRUE)) %>% ungroup()
# Editing chmistry geoMean by substituting "." (below detection limit)  into t1/3 of min of that chemistry 
contaminants.coldata.subset <- read_csv("GLRI_TRES_contaminantsto2020_Transcriptome_subset_CT.csv")
contaminants.coldata.subset <- contaminants.coldata.subset %>% filter(!is.na(Value)) ## remove those with only NA
contaminants.coldata.subset <- contaminants.coldata.subset %>% left_join(contaminants.min.adjusted, by = c("Key","class"))
contaminants.coldata.subset <- contaminants.coldata.subset %>% filter(!minValue == "Inf") ## remove those with only "."
contaminants.coldata.subset$value.adjust[contaminants.coldata.subset$Value == "."] <- contaminants.coldata.subset$minValue[contaminants.coldata.subset$Value == "."]/3 # "." converts to 1/3 of loweast detetable value
# Load Gene ID MAP for GeneID to GeneName conversion
ID.convert <- readRDS("IDmap_final.rds") ## gene ID Convert table
siteMAP.all <- as_tibble(readRDS("siteMAP_all.rds")) ## site ID convert Table
siteMAP.all$SiteID[siteMAP.all$propersite == "StarLake"] <- "SL"
dds.bothSex.adjusted <- readRDS("dds.bothSex.adjusted.outlierRM.rds") ## normalized count matrix
dds.bothSex.adjusted <- dds.bothSex.adjusted[,colnames(dds.bothSex.adjusted) != "19HW576B"] ## remove "19HW576B" only use A chick for the analysis for regression analysis
colnames(dds.bothSex.adjusted) <- gsub("A$|B$", "", colnames(dds.bothSex.adjusted)) ## remove all the A or B tail 
dds.bothSex.adjusted$propersite2 <- as.character(siteMAP.all$propersite[match(dds.bothSex.adjusted$site, siteMAP.all$SiteID)]) 
counts.tswallow <- assay(dds.bothSex.adjusted)
Cont.major.list <- c("Nonachlor, cis-_C", "Heptachlor Epoxide_C",   "Nonachlor, trans-_C",  "PFDA_C", "Chlordane, oxy-_C", "Total Parent PAHs", "Total PBDE", "Total PAHs","Total PCBs","Total_PFCs")
## contaminants.coldata.subset2: There are some PCBs are LRPCBs, some are HRPCBs. Combining all of  "TOTAL PCBs_C", and "TOTAL PCBs" into "TOTAL PCBs" and if there are overlapping, pick LRPCBs first 
contaminants.coldata.subset <- contaminants.coldata.subset %>% filter(class != "PAHs")
contaminants.coldata.subset <- contaminants.coldata.subset %>% mutate(MergeSite=replace(MergeSite, MergeSite == "ref", "SL")) ## replace ref to SL 
contaminants.coldata.subset$MergeID <- gsub("-","", contaminants.coldata.subset$MergeID) ## remove all dash
contaminants.coldata.subset$MergeID <- gsub("A$","", contaminants.coldata.subset$MergeID) ## remove A tail
contaminants.coldata.subset <- contaminants.coldata.subset %>% mutate(Key=replace(Key, Key == "TOTAL PCBs_C" | Key == "TOTAL PCBs", "TOTAL PCBs"))
counts.tswallow.norm <- vst(dds.bothSex.adjusted)
counts.tswallow.norm <- assay(counts.tswallow.norm)
```
#### Load self-defined functions for PCB training

GetTopDEGs_byGeoMean.byNest(Con,Con_ind)
: get DEGs (p < 0.05) and select top 1000 most differentiating genes using EdgeR against PCBs tissue concentrations  

Run_lasso.onesiteout(n, Con, Con_ind, s, lam.m)
: Run leave-one-out (by site) cross validation and train lasso regression against PCBs tissue concentrations; 31 sites, sample n out of  31*300 re-sampling  


```{r Load self-defined functions for PCB training, eval = FALSE}
GetTopDEGs_byGeoMean.byNest <- function(Con,Con_ind) {
  nestid <- contaminants.coldata.subset2 %>% filter(Key == Con_ind) %>% dplyr::pull(MergeID)
  match1 <- colnames(counts.tswallow) %in% nestid
  counts.con  <-  counts.tswallow[,match1]
  counts.con.batch <- as.factor(as.character(dds.bothSex.adjusted$batch2)[match1])
  counts.con.sex <- as.factor(as.character(dds.bothSex.adjusted$sex)[match1])
  contamin.matrix <- contaminants.coldata.subset2 %>% dplyr::slice(match(colnames(counts.con), contaminants.coldata.subset2$MergeID)) %>% filter(Key == Con_ind)
  counts.con.contamin <- log10(contamin.matrix$value.adjust) ## get log10 value to suppress potential outlier
  ## EdgeR needs non-normalized counts
  y <- DGEList(counts=as.matrix(counts.con))
  y <- calcNormFactors(y)
  if (length(levels(counts.con.batch)) > 1) {
    design <- model.matrix(~counts.con.batch+counts.con.sex+counts.con.contamin) ## add batch as cofactor
    y <- estimateDisp(y,design)
    fit <- glmQLFit(y,design)
    qlf <- glmQLFTest(fit, coef=4)} else {
      design <- model.matrix(~counts.con.sex+counts.con.contamin) ## add batch as cofactor
      y <- estimateDisp(y,design)
      fit <- glmQLFit(y,design) 
      qlf <- glmQLFTest(fit, coef=3)
    }
  DEGs.Q <- sum(decideTests(qlf) != 0)
  print(DEGs.Q)
  match2 <- row.names(topTags(qlf, n=1000))
  match2.logFC <- topTags(qlf, n=1000)[[1]][,"logFC"]
  match2.table <- topTags(qlf, n=2000)
  match2.table.all <- topTags(qlf, n=nrow(qlf))
  match2.DEGs.table <- topTags(qlf, n=DEGs.Q)
  assign(paste0("DEGs.table_by_GeoMeanNest.",Con,Con_ind),match2.DEGs.table, envir = parent.frame())
  assign(paste0("top1000DEGs_by_GeoMeanNest.",Con,Con_ind),match2, envir = parent.frame())
  assign(paste0("top1000DEGs_by_GeoMeanNest.direction",Con,Con_ind),match2.logFC, envir = parent.frame())
  assign(paste0("DEGs.top2000.table_by_GeoMeanNest.",Con,Con_ind),match2.table, envir = parent.frame())
  assign(paste0("DEGs.allGene.table_by_GeoMeanNest.",Con_ind),match2.table.all, envir = parent.frame())
} # get DEGs (p < 0.05) and select top 1000 most differentiating genes using EdgeR against PCBs tissue concentrations

assessmargin <- function(accuracy.Q.accum){
  qt(0.975,df=length(accuracy.Q.accum)-1)*sd(accuracy.Q.accum)/sqrt(length(accuracy.Q.accum))
} # determine margin

lasso.by.nest.onesiteout.test <- function(Con,Con_ind,i) {
  
  nestid <- contaminants.coldata.subset2 %>% filter(Key == Con_ind) %>% dplyr::pull(MergeID)
  match1 <- colnames(counts.tswallow.norm) %in% nestid
  counts.con  <-  counts.tswallow.norm[,match1]
  counts.con.batch <- as.factor(as.character(dds.bothSex.adjusted$batch2)[match1])
  counts.con.sex <- as.factor(as.character(dds.bothSex.adjusted$sex)[match1])
  contamin.raw <- contaminants.coldata.subset2 %>% filter(Key == Con_ind) %>% dplyr::pull(value.adjust) %>% log10()
  contamin.matrix <- contaminants.coldata.subset2 %>% dplyr::slice(match(colnames(counts.con), contaminants.coldata.subset2$MergeID)) %>% filter(Key == Con_ind)
  counts.con.contamin <- log10(contamin.matrix$value.adjust) ## get log10 value to suppress potential outlier
  ## Building matrix 
  if (length(levels(counts.con.batch)) > 1) {
    counts.con.t <- as.data.frame(t(counts.con[get(paste0("top1000DEGs_by_GeoMeanNest.",Con,Con_ind)),]))
    counts.con.t <- cbind(counts.con.t, batch = counts.con.batch, sex = counts.con.sex, contaminant = counts.con.contamin)
    m <- model.matrix(contaminant ~ batch +., counts.con.t)
    m <- m[,-1] } else {
      counts.con.t <- as.data.frame(t(counts.con[get(paste0("top1000DEGs_by_GeoMeanNest.",Con,Con_ind)),]))
      counts.con.t <- cbind(counts.con.t, sex = counts.con.sex, contaminant = counts.con.contamin)
      m <- model.matrix(contaminant ~ ., counts.con.t)
      m <- m[,-1] }  
  ## make sure most of sites have testing samples, allsite: # of all genomic individuals, nestsite: # of genomic individuals with matching nest chemical info
  counts.site  <-  dds.bothSex.adjusted$propersite2[match1]
  matchsubset.train <- counts.site != unique(counts.site)[i] # train subset
  print(c("training #", sum(matchsubset.train)))
  matchsubset.test <- counts.site != unique(counts.site)[i]
  l.lse <- {}
  for (j in 1:20) {
    cvfit <- cv.glmnet(x=m[matchsubset.train,], y=counts.con.t$contaminant[matchsubset.train], nfold = 10, alpha = 1, lambda = 10^seq(0,-2,length=200), relax =FALSE)
    l.lse = c(l.lse, cvfit$lambda.1se)
  }
  l.lse = median(l.lse)
  cvfit1 <- cv.glmnet(x=m[matchsubset.train,], y=counts.con.t$contaminant[matchsubset.train], nfold = 10, alpha = 1, lambda = 10^seq(0,-2,length=600), relax =FALSE)
  ## only plot it 3 times in test runs
  pdf(paste0("~/Documents/work/Tswallow_chem_GLRI_update/Temp_plots/ByNest_",Con_ind, "_distribution",".pdf"))
  plot(sort(counts.con.contamin),main = paste0("ByNest.",Con_ind, ".distribution"), ylab = "log10(value)")
  dev.off()
  
  pdf(paste0("~/Documents/work/Tswallow_chem_GLRI_update/Temp_plots/lamdaPlot_byNest_onesiteout",Con_ind, ".pdf"))
  plot(cvfit1,main = paste0("R_lamdaPlot_leaveonesiteout.byNest.",Con_ind))
  
  dev.off()
  } # test Run 

## Run leave-one-out (by site) cross validation and train lasso regression against PCBs tissue concentrations 
lasso.by.nest.onesiteout <- function(Con,Con_ind, s, lam.m,i) {
  nestid <- contaminants.coldata.subset2 %>% filter(Key == Con_ind) %>% dplyr::pull(MergeID)
  match1 <- colnames(counts.tswallow.norm) %in% nestid
  counts.con  <-  counts.tswallow.norm[,match1]
  counts.con.batch <- as.factor(as.character(dds.bothSex.adjusted$batch2)[match1])
  counts.con.sex <- as.factor(as.character(dds.bothSex.adjusted$sex)[match1])
  contamin.raw <- contaminants.coldata.subset2 %>% filter(Key == Con_ind) %>% dplyr::pull(value.adjust) %>% log10()
  contamin.matrix <- contaminants.coldata.subset2 %>% dplyr::slice(match(colnames(counts.con), contaminants.coldata.subset2$MergeID)) %>% filter(Key == Con_ind)
  counts.con.contamin <- log10(contamin.matrix$value.adjust) ## get log10 value to suppress potential outlier
  ## Building matrix 
  if (length(levels(counts.con.batch)) > 1) {
    counts.con.t <- as.data.frame(t(counts.con[get(paste0("top1000DEGs_by_GeoMeanNest.",Con,Con_ind)),]))
    counts.con.t <- cbind(counts.con.t, batch = counts.con.batch, sex = counts.con.sex, contaminant = counts.con.contamin)
    m <- model.matrix(contaminant ~ batch +., counts.con.t)
    m <- m[,-1] } else {
      counts.con.t <- as.data.frame(t(counts.con[get(paste0("top1000DEGs_by_GeoMeanNest.",Con,Con_ind)),]))
      counts.con.t <- cbind(counts.con.t, sex = counts.con.sex, contaminant = counts.con.contamin)
      m <- model.matrix(contaminant ~ ., counts.con.t)
      m <- m[,-1] }  
  ## make sure most of sites have testing samples, allsite: # of all genomic individuals, nestsite: # of genomic individuals with matching nest chemical info
  counts.site  <-  dds.bothSex.adjusted$propersite2[match1]
  matchsubset.train <- counts.site != unique(counts.site)[i] # train subset
  ## sample 90% 
  matchsubset.train <- sample(which(matchsubset.train),(sum(matchsubset.train)*0.9))
  print(c("training #", length(matchsubset.train)))
  matchsubset.test <- counts.site == unique(counts.site)[i]
  l.1se <- {}
  l.min <- {}
  for (j in 1:20) {
    cvfit <- cv.glmnet(x=m[matchsubset.train,], y=counts.con.t$contaminant[matchsubset.train], nfold = 10, alpha = 1, lambda = 10^seq(0,-2,length=200), relax =FALSE)
    l.1se <- c(l.1se, cvfit$lambda.1se)
    l.min <- c(l.min, cvfit$lambda.min)
  }
  l.1se = median(l.1se)
  l.min <- median(l.min)
  l.median <- exp(mean(c(log(l.min),log(l.1se))))
  lamda.sequence <- c(l.1se,l.min,l.median)
  ByNestResult.withlamdaP <- {}
  
  fit <- glmnet(x=m[matchsubset.train,], y=counts.con.t$contaminant[matchsubset.train], alpha = 1, lambda = 10^seq(0,lam.m,length=600), relax = FALSE)
  Result <- assess.glmnet(fit, newx = m[matchsubset.train,], newy = counts.con.t$contaminant[matchsubset.train], s = lamda.sequence[s])
  
  Result.lasso.assessment <- {}
  Result.lasso.assessment$mse <- Result$mse[[1]]
  Result.lasso.assessment$mae <- Result$mae[[1]]
  
  variables <- coef(fit, s = lamda.sequence[s])
  Result.lasso.assessment$variables <- row.names(variables)[!(variables[,1] == 0)][-1] 
  pseudo_R2 <- fit$dev.ratio[which(sort(c(lamda.sequence[s],10^seq(0,lam.m,length=600)),decreasing = TRUE) == lamda.sequence[s])[1] -1]
  Result.lasso.assessment$pseudo_R2 <- pseudo_R2
  
  ## Building test matrix 
  predict.site.test <- predict(fit, newx = m[matchsubset.test,], s = lamda.sequence[s]) 
  result.list = list(lamda.sequence,Result.lasso.assessment,predict.site.test)
  names(result.list) <- c("lamda.sequence","Result.relax.lasso.assessment","predictions")
  return(result.list)
} 
Run_lasso.onesiteout <- function(n, Con, Con_ind, s, lam.m) {
    lassoNestResult.run <- foreach(i = sample(rep(1:31,300),n), .packages = c("dplyr","glmnet"), .export = c("lasso.by.nest.onesiteout","contaminants.coldata.subset2","counts.tswallow.norm","dds.bothSex.adjusted", paste0("top1000DEGs_by_GeoMeanNest.",Con,Con_ind), "data.contaminants.majorSubset.geo.bysite")) %dopar% {
    r_lasso_result <- Vectorize(lasso.by.nest.onesiteout)(Con,Con_ind, s, lam.m,i)
    return(r_lasso_result)
  }
  assign(paste0("lassoResult.onesiteout",sub(" ","",Con_ind)), lassoNestResult.run, envir = parent.frame())
  return(lassoNestResult.run)
} # 31 sites, sample n out of  31*300 re-sampling 
```

#### Lasso regression analysis between global geene expression and PCB tissue concentrations
Run lasso with leave one (site) cross validation and by selecting 1000 individuals from (31*300 resampling = 9300), using 1.1se   

topgenelistPCBs.rds
: selecting 91 genes which appeared more than 50 times (> 5%) in the cross-validation

```{r Train PCB lasso model and cross-validaiton, eval = FALSE}
## Total PCBs
Con="PCBs_data";Con_ind="TOTAL PCBs"
## refresh between chemicals 
lamda.sequence <- {}
Result.lasso.assessment <- {}
variables <- {}
predict.site.test <- {}
rlassoResult <- list()
## Set up PCBs chemistry data 
contaminants.coldata.subset2 <-contaminants.coldata.subset
## Individual chemicals 
contaminants.coldata.subset2 <- contaminants.coldata.subset2 %>% filter(Key == Con_ind)
contaminants.coldata.subset2 <- contaminants.coldata.subset2 %>% filter(!is.na(value.adjust)) ## remove those with value == "NR"
## combine all duplicated items using mean value because they have the same nest id 
contaminants.coldata.subset2 <- contaminants.coldata.subset2 %>% group_by(MergeID,Species,Matrix,AOC,Proper_Site,Key,type,class,MergeSite,proper_site2) %>% summarise(value.adjust2 = mean(value.adjust)) %>% ungroup()
colnames(contaminants.coldata.subset2)[colnames(contaminants.coldata.subset2) == "value.adjust2"] <- "value.adjust" ## change the name back
print(sum(duplicated(contaminants.coldata.subset2$MergeID))) ## check duplication again
TotalPCBs.mergeIDlist = contaminants.coldata.subset2$MergeID  
## Get top DEGs using DESeq2 against PCBs gradient  
GetTopDEGs_byGeoMean.byNest(Con=Con,Con_ind=Con_ind) #103 DEGs (p < 0.05)
# saveRDS(`DEGs.allGene.table_by_GeoMeanNest.TOTAL PCBs`, "~/Documents/work/Tswallow_chem_GLRI_update/GLRI_MS_figures/TOP.PCBs.allgene.table.bynest.rds")
## Train lasso regression model using glmnet; relax lasso was not included because there was no significant improvement  

## Register parallel 
cl <- parallel::makeCluster(12)
doParallel::registerDoParallel(cl)
lassoResult.totalPCBs = Run_lasso.onesiteout(n= 1000, Con=Con,Con_ind=Con_ind, s= 1, lam.m=-2) # lamda.sequence <- c(l.1se,l.min,l.median); run lasso with leave one (site) cross validation and by selecting 1000 individuals from (31*300 resampling = 9300), using 1.1se 
parallel::stopCluster(cl) # stop parallel
## selecting those genes which were selected more than 50 times (> 5%)
test.genelist.topPCB <- names(table(unlist(sapply(1:1000, function(x) lassoResult.totalPCBs[[x]][[2]]$variables))))[table(unlist(sapply(1:1000, function(x) lassoResult.totalPCBs[[x]][[2]]$variables))) > 50]
## save top gene 
saveRDS(test.genelist.topPCB, file ="topgenelistPCBs.rds") # 91 genes  

```

## 2. For selecting top predictor genes to predcit PAHs (by pooled stomach contents each site)  

### Process coldata by site 
```{r Process chemistry by site, eval = FALSE}
data.contaminants.majorSubset.geo.bysite$MergeSite[data.contaminants.majorSubset.geo.bysite$MergeSite == "ref"] <- "SL" ## convert ref to SL 
## combine all duplicated items by MergeSite
data.contaminants.majorSubset.geo.bysite <- data.contaminants.majorSubset.geo.bysite %>% group_by(proper_site2,MergeSite,Key) %>% summarise(value.geoMean2 = mean(value.geoMean)) %>% ungroup()
colnames(data.contaminants.majorSubset.geo.bysite)[colnames(data.contaminants.majorSubset.geo.bysite) == "value.geoMean2"] <- "value.geoMean" ## change the name back
```

### Load self-defined functions for lasso regression training
GetTopDEGs_byGeoMean.bySiteGeoMean()
: get top DEGs using edgeR linear regression against PAHs concentrations  

lasso.by.site()
: using top1000 DEGs for lasso regression analysis with 10-fold validation to determine lamda, leave one (site) out cross-validation, resample (sample 90%) of training data  

Run_lasso_by_site
: Run n runs of lasso.by.site()  


```{r self-defined functions for PAHs lasso regression analysis, eval = FALSE}
GetTopDEGs_byGeoMean.bySiteGeoMean <- function(Con,Con_ind) {
  
  match1 <- str_sub(colnames(counts.tswallow),3,4) %in% contaminants.coldata.subset2$MergeSite
  counts.con  <-  counts.tswallow[,match1]
  counts.con.batch <- as.factor(as.character(dds.bothSex.adjusted$batch2)[match1])
  counts.con.sex <- as.factor(as.character(dds.bothSex.adjusted$sex)[match1])
  counts.con.contamin <- log10(contaminants.coldata.subset2$value.geoMean[match(str_sub(colnames(counts.con),3,4), contaminants.coldata.subset2$MergeSite)])
  
  ## EdgeR needs non-normalized counts
  y <- DGEList(counts=as.matrix(counts.con))
  y <- calcNormFactors(y)
  if (length(levels(counts.con.batch)) > 1) {
    design <- model.matrix(~counts.con.batch+counts.con.sex+counts.con.contamin) ## add batch as cofactor
    y <- estimateDisp(y,design)
    fit <- glmQLFit(y,design)
    qlf <- glmQLFTest(fit, coef=4)} else {
      design <- model.matrix(~counts.con.sex+counts.con.contamin) ## add batch as cofactor
      y <- estimateDisp(y,design)
      fit <- glmQLFit(y,design) 
      qlf <- glmQLFTest(fit, coef=3)
    }
  DEGs.Q <- sum(decideTests(qlf) != 0)
  print(DEGs.Q)
  match2 <- row.names(topTags(qlf, n=1000))
  match2.logFC <- topTags(qlf, n=1000)[[1]][,"logFC"]
  match2.table <- topTags(qlf, n=2000)
  match2.table.all <- topTags(qlf, n=nrow(qlf))
  match2.DEGs.table <- topTags(qlf, n=DEGs.Q)
  assign(paste0("DEGs.table_by_GeoMeanSite.",Con,Con_ind),match2.DEGs.table, envir = parent.frame())
  assign(paste0("top1000DEGs_by_GeoMeanSite.",Con,Con_ind),match2, envir = parent.frame())
  assign(paste0("top1000DEGs_by_GeoMeanSite.direction",Con,Con_ind),match2.logFC, envir = parent.frame())
  assign(paste0("DEGs.top2000.table_by_GeoMeanSite.",Con,Con_ind),match2.table, envir = parent.frame())
  assign(paste0("DEGs.allGene.table_by_GeoMeanSite.",Con_ind),match2.table.all, envir = parent.frame())
}

# Min_L: for how low the lamda cv goes 10^(-2,-3,-4); s: lamda.sequence: 1) 1se 2) min 3) log median 
lasso.by.site <- function(Con,Con_ind,i,Min_L,s) {
  BySiteResult.withlamdaP <- {} ## refresh
  variables.all <- list()
  predict.site.test.all <- list()
  match1 <- str_sub(colnames(counts.tswallow),3,4) %in% contaminants.coldata.subset2$MergeSite
  counts.con  <-  counts.tswallow.norm[,match1]
  counts.con.batch <- as.factor(as.character(dds.bothSex.adjusted$batch2)[match1])
  counts.con.sex <- as.factor(as.character(dds.bothSex.adjusted$sex)[match1])
  counts.con.contamin <- log10(contaminants.coldata.subset2$value.geoMean[match(str_sub(colnames(counts.con),3,4), contaminants.coldata.subset2$MergeSite)])
  ## Building matrix 
  if (length(levels(counts.con.batch)) > 1) {
    counts.con.t <- as.data.frame(t(counts.con[get(paste0("top1000DEGs_by_GeoMeanSite.",Con,Con_ind)),]))
    counts.con.t <- cbind(counts.con.t, batch = counts.con.batch, sex = counts.con.sex, contaminant = counts.con.contamin)
    m <- model.matrix(contaminant ~ batch +., counts.con.t)
    m <- m[,-1] } else {
      counts.con.t <- as.data.frame(t(counts.con[get(paste0("top1000DEGs_by_GeoMeanSite.",Con,Con_ind)),]))
      counts.con.t <- cbind(counts.con.t, sex = counts.con.sex, contaminant = counts.con.contamin)
      m <- model.matrix(contaminant ~ ., counts.con.t)
      m <- m[,-1] }  
  
  ## determinne training and testing set
  counts.site  <-  dds.bothSex.adjusted$propersite2[match1]
  match1.test <- counts.site == unique(counts.site)[i]
  match1.test.sample <- sample(which(match1.test), size = median(table(counts.site)), replace = TRUE)
  matchsubset.train <- counts.site != unique(counts.site)[i] # train subset
  ## sample 90% of training set
  matchsubset.train <- sample(which(matchsubset.train),(sum(matchsubset.train)*0.9))
  
  l.1se <- {} ## repeat 5 times to avoid outlier in cross validation 
  l.min <- {}
  # foldid<- as.numeric(as.factor(str_sub(colnames(counts.con[,matchsubset.train]),3,4))) # don't use foldid as batch effect for consistency

  cvfit <- cv.glmnet(x=m[matchsubset.train,], y=counts.con.t$contaminant[matchsubset.train], nfolds = 10, alpha = 1, lambda = 10^seq(0,Min_L,length=600))
  l.1se <- cvfit$lambda.1se
  l.min <- cvfit$lambda.min
  l.median <- exp(mean(c(log(l.1se),log(l.min))))
  lamda.sequence <- c(l.1se,l.min,l.median)  
  names(lamda.sequence) <- c("1se","min","median")
    fit <- glmnet(x=m[matchsubset.train,], y=counts.con.t$contaminant[matchsubset.train], alpha = 1, lambda = 10^seq(0,Min_L,length=600))
    Result1 <- assess.glmnet(fit, newx = m[match1.test.sample,], newy = counts.con.t$contaminant[match1.test.sample], s = lamda.sequence[s]) 
    BySiteResult.withlamdaP$mse <- Result1$mse
    BySiteResult.withlamdaP$mae <- Result1$mae
    variables <- coef(fit, s = lamda.sequence[s])  
    variables <- row.names(variables)[!(variables[,1] == 0)]
    variables <- variables[-1]
    
    pseudo_R2 <- fit$dev.ratio[which(sort(c(lamda.sequence[s],10^seq(0,Min_L,length=600)),decreasing = TRUE) == lamda.sequence[s])[1] -1]
    predict.site.test1 <- predict(fit, newx = m[match1.test.sample,], s = lamda.sequence[s]) 
    predict.site.test.all <- predict.site.test1
    
    result = list(lamda.sequence,BySiteResult.withlamdaP,variables,pseudo_R2,predict.site.test.all)
    names(result) <- c("lamda.sequence","mse_mae","variables","pseudo_R2","predict.site")
    return(result)
}
Run_lasso_by_site <- function(n,Con,Con_ind,Min_L,s) {
TIME1 <- Sys.time()
lassositeResult.run <- foreach(i = rep(1:28,n), .packages = c("dplyr","glmnet"), .export = c("lasso.by.site","contaminants.coldata.subset2","counts.tswallow.norm","dds.bothSex.adjusted", paste0("top1000DEGs_by_GeoMeanSite.",Con,Con_ind), "data.contaminants.majorSubset.geo.bysite")) %dopar% {
  cat(paste("Starting iteration",i,Sys.time(),"\n"), file = paste0("~/Documents/work/Tswallow_chem_GLRI_update/relax_lasso_result/",Con_ind,"lassoBySiteRun.log2.txt"), append = TRUE)
  lasso_result <- Vectorize(lasso.by.site)(Con,Con_ind,i,Min_L,s)
  return(lasso_result)
}
return(lassositeResult.run)
TIME2 <- Sys.time()
print(TIME2 - TIME1)
} 
```

### Lasso regression analysis between global geene expression and PAH concentrations
Run lasso with leave one (site) cross validation, lamda chose l.1se for selecting top predictor genes from top 1000 genes (EdgeR, linear regression against PAHs concentrations)
lasso_topgene_bySite_PAHs.rds
: selecting top 110 genes in the cross-validation (> 10% of cross-validation)  
```{r Train PAH lasso model and cross-validaiton, eval = FALSE}

Con="PAHs_data";Con_ind="Total PAHs"
## edit subset for overlapping genomic samples with contaminant data 
contaminants.coldata.subset2 <-data.contaminants.majorSubset.geo.bysite
## Individual chemicals 
contaminants.coldata.subset2 <- contaminants.coldata.subset2 %>% filter(Key == Con_ind)
## combine all duplicated items using mean value because they have the same MergeSite ## only River Raisin will be affected 
contaminants.coldata.subset2 <- contaminants.coldata.subset2 %>% group_by(MergeSite, Key) %>% summarise(value.geoMean2 = mean(value.geoMean)) %>% ungroup()
colnames(contaminants.coldata.subset2)[colnames(contaminants.coldata.subset2) == "value.geoMean2"] <- "value.geoMean" ## change the name back to value.geoMean
print(sum(duplicated(contaminants.coldata.subset2$MergeSite))) ## check duplication again
GetTopDEGs_byGeoMean.bySiteGeoMean(Con = Con, Con_ind = Con_ind) ## 570 DEGs (p < 0..05) against PAHs tissue concentrations
# saveRDS(`DEGs.allGene.table_by_GeoMeanSite.Total PAHs`,"~/Documents/work/Tswallow_chem_GLRI_update/GLRI_MS_figures/TOP.PAHs.allgene.table.bysite.rds")

## Run lasso regression by site to predict PAHs tissue concentrations
registerDoParallel(cores = 12)
## Becasue the purpose 
lassoResult.bySite.TotalPAHs = Run_lasso_by_site(n=200,Con=Con,Con_ind = Con_ind, Min_L = -2, s = 1) # lamda.sequence <- c(l.1se,l.min,l.median)
parallel::stopCluster(cl) # stop parallel
lassoResult.bySite = lassoResult.bySite.TotalPAHs
# saveRDS(lassoResult.bySite.TotalPAHs,"~/Documents/work/Tswallow_chem_GLRI_update/GLRI_MS2_all/lasso_topgene_bySite_PAHs.rds")

## 100% trials have less than 95 genes, go ahead and pick genes appears at 10% of trials 
lasso_selected_topgene <- names(sort(table(unlist(sapply(1:(28*200),function(x) lassoResult.bySite[[x]][[3]]))),decreasing = TRUE)[1:110])
# saveRDS(lasso_selected_topgene, "~/Documents/work/Tswallow_chem_GLRI_update/GLRI_MS2_all/lasso_topgene_bySite_PAHs.rds")
```

